Reference Tracking Strategies of Deaf Adult Signers in Turkish Sign Language

### 2023-05-04 revised_manuscript.Rmd
update

```{r echo=FALSE, message=FALSE, cache=FALSE, results='asis', warning=FALSE}
colnames(df_model_1)[1:4] <- c("Coefficients","outerWidth","innerWidth", "pointEstimate")
df_model_1 %<>% mutate(across(.cols = "Coefficients", ~str_replace( ., "_","\\\\_")))
apa_table(df_model_1,
          format = "latex",
          booktabs = T,
          escape = FALSE,
          landscape = FALSE,
          font_size = "small",
          placement = "hbt!",
          caption ="Regression model results for the RE Type (Nominal, Zero, Classifier) with the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (DoD, DoH)"
)
```

>>>>

```{r echo=FALSE, message=FALSE, cache=FALSE, results='asis', warning=FALSE}
colnames(df_model_1)[1:4] <- c("Coefficients","outerWidth","innerWidth", "pointEstimate")
df_model_1 %<>% dplyr::select(-model) %>% mutate(across(.cols = "Coefficients", ~str_replace( ., "_","\\\\_")))
apa_table(df_model_1,
          format = "latex",
          booktabs = T,
          escape = FALSE,
          landscape = FALSE,
          font_size = "small",
          placement = "hbt!",
          caption ="Regression model results for the RE Type (Nominal, Zero, Classifier) with the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (DoD, DoH)"
)
```

### 2023-05-04 revised_manuscript.tex
update

% Pandoc header
\usepackage{floatrow}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{float}
\floatplacement{figure}{H}
\floatplacement{table}{H}
\floatsetup[figure]{capposition=top}
\floatsetup[table]{capposition=top}
\usepackage{times}
\usepackage{smartdiagram}
\usepackage{ragged2e}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{./sty/langsci-gb4e}
\usepackage{caption}
\usepackage{multicol}
\setlength{\columnsep}{-2cm}
\usepackage{setspace}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

Sign languages are the natural languages of the Deaf communities all
around the world. They are independent languages, which do not derive
from majority spoken language(s) but emerge naturally when deaf people
form a community. Turkish Sign Language (Türk İşaret Dili -- TİD) is and
has been the officially recognized language of the Deaf community in
Turkey since 2005 (see Nuhbalaoglu, 2018 for a brief review of the
historical development of the language). Deaf or hearing children
acquiring a SL within a family where there is one or more deaf
caregivers display the same developmental milestones as hearing children
acquiring a spoken language in a hearing family (Chamberlain et al.,
1999; Meier and Newport, 1990). However, reports indicate that over 90\%
of all deaf individuals are born into hearing families (DoH) who do not
know how to sign (Woll, 2013). This means that the auditory modality
used in the household is not accessible to many deaf children. That is
why the age of acquisition of a first functional language varies greatly
among deaf children, with many of whom being deprived of their first
languages typically until after age 7 when they start formal education
in a deaf school, especially in the context of Turkey (Sari, 2005).
Despite this, many of the early intervention programs for late signing
children prioritize spoken language development and discourage the use
of a sign language (Kemaloğlu and Kemaloğlu, 2012). The Ministry of
National Education (2020) identifies 72 deaf schools in Turkey for the
K-12 level. Even when deaf children with hearing parents start their
formal education in one of these schools, they do not receive proper
instruction in TİD since the teachers themselves are mainly hearing and
only have a very rudimentary knowledge of a sign language (Ilkbasaran,
2015; Sari, 2005). In some cases, it has been argued that late exposure
to a sign language among certain DoH signers has negative effects on
some cognitive and linguistic measures. To name a few, delayed exposure
to a first functional language is reported to negatively affect deaf
children and adults' morphosyntactic production and processing
(Boudreault and Mayberry, 2006; Cheng and Mayberry, 2021; Cormier et
al., 2012; Karadöller et al., 2017; Kayabaşı and Gökgöz, 2022; Newport,
1990) and narrative abilities (e.g., Becker, 2009). In our overall
research, we are concerned with the effects of this late exposure in the
ultimate adult forms of TİD users.

@@

For more inference on the data and having pairwise comparisons, we fit a
regression model to RE Type using discourse status and acquisition group
as predictors. We used the brms package (Bürkner, 2018) in R, with a
categorical distribution on the responses and sum contrasts for the
predictors Discourse Status (Introduction, Maintenance, Re-introduction)
and Participant Group (DoD, DoH). As the overall number of produced
pronouns was very little (40 data points), we excluded pronominal use
from the following analyses. For the current analysis, we did not define
priors ourselves and used the preset uninformed priors. Figure
2\footnote{We ran another model where we used the self-reported
  acquisition time as a predictor with 3 levels 0-3:Early, 4-7:Late1,
  and 7-above:Late2. However, our model did not converge properly and
  the credible intervals were very wide spread, yielding the results
  uninformative. We provide the model results as a plot in
  \ref{section:d}.} shows the posterior probability distributions where
the dot is the median estimate, the thick line is the 50\% credible
intervals, and the thin line is the 95\% credible intervals. There are 3
responses we are considering in this model. They are Classifier,
Nominal, and Zero Anaphora. Classifier response is the level that other
responses Nominal and Zero Anaphora are compared against.

@@

@@

\caption{\label{tab:unnamed-chunk-4}Model results for Nativeness with 3 levels}

@@

\leavevmode\vadjust pre{\hypertarget{ref-chamberlain_language_1999}{}}%
Chamberlain, Charlene, Morford, Jill P., Mayberry, Rachel I. (Eds.),
1999. Language acquisition by eye. Psychology Press, Mahwah, N.J.

@@

\leavevmode\vadjust pre{\hypertarget{ref-czubek2017}{}}%
Czubek, Todd Alan, 2017. A comprehensive study of referring expressions
in ASL 244.

@@

\leavevmode\vadjust pre{\hypertarget{ref-engberg-pedersen1993}{}}%
Engberg-Pedersen, Elisabeth, 1993. Space in danish sign language: The
semantics and morphosyntax of the use of space in a visual language.

@@

\leavevmode\vadjust pre{\hypertarget{ref-ministryofnationaleducation2020}{}}%
Ministry of National Education, 2020.
\href{http://sgb.meb.gov.tr/meb_iys_dosyalar/2020_09/04144812_meb_istatistikleri_orgun_\%20egitim_2019_2020.pdf}{Milli
e{ğ}itim istatistikleri, örgün e{ğ}itim 2019/{'}20 {[}national education
statistics, formal education 2019/{'}20{]}}.

@@

>>>>

% Pandoc header
\usepackage{floatrow}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{float}
\floatplacement{figure}{H}
\floatplacement{table}{H}
\floatsetup[figure]{capposition=top}
\floatsetup[table]{capposition=top}
\usepackage{times}
\usepackage{smartdiagram}
\usepackage{ragged2e}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{langsci-gb4e}
\usepackage{caption}
\usepackage{multicol}
\setlength{\columnsep}{-2cm}
\usepackage{setspace}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

Sign languages are the natural languages of the Deaf communities all
around the world. They are independent languages, which do not derive
from majority spoken language(s) but emerge naturally when deaf people
form a community. Turkish Sign Language (Türk İşaret Dili -- TİD) is and
has been the officially recognized language of the Deaf community in
Turkey since 2005 (see Nuhbalaoglu, 2018 for a brief review of the
historical development of the language). Deaf or hearing children
acquiring a SL within a family where there is one or more deaf
caregivers display the same developmental milestones as hearing children
acquiring a spoken language in a hearing family (Chamberlain et al.,
1999; Meier and Newport, 1990). However, reports indicate that over 90\%
of all deaf individuals are born into hearing families (DoH) who do not
know how to sign (Woll, 2013). This means that the auditory modality
used in the household is not accessible to many deaf children. That is
why the age of acquisition of a first functional language varies greatly
among deaf children, with many of whom being deprived of their first
languages typically until after age 7 when they start formal education
in a deaf school, especially in the context of Turkey (Sari, 2005).
Despite this, many of the early intervention programs for late signing
children prioritize spoken language development and discourage the use
of a sign language (Kemaloğlu and Kemaloğlu, 2012). The National
Education (2020) identifies 72 deaf schools in Turkey for the K-12
level. Even when deaf children with hearing parents start their formal
education in one of these schools, they do not receive proper
instruction in TİD since the teachers themselves are mainly hearing and
only have a very rudimentary knowledge of a sign language (Ilkbasaran,
2015; Sari, 2005). In some cases, it has been argued that late exposure
to a sign language among certain DoH signers has negative effects on
some cognitive and linguistic measures. To name a few, delayed exposure
to a first functional language is reported to negatively affect deaf
children and adults' morphosyntactic production and processing
(Boudreault and Mayberry, 2006; Cheng and Mayberry, 2021; Cormier et
al., 2012; Karadöller et al., 2017; Kayabaşı and Gökgöz, 2022; Newport,
1990) and narrative abilities (e.g., Becker, 2009). In our overall
research, we are concerned with the effects of this late exposure in the
ultimate adult forms of TİD users.

@@

For more inference on the data and having pairwise comparisons, we fit a
regression model to RE Type using discourse status and acquisition group
as predictors. We used the brms package (Bürkner, 2018) in R, with a
categorical distribution on the responses and sum contrasts for the
predictors Discourse Status (Introduction, Maintenance, Re-introduction)
and Participant Group (DoD, DoH). As the overall number of produced
pronouns was very little (40 data points), we excluded pronominal use
from the following analyses. For the current analysis, we did not define
priors ourselves and used the preset uninformed priors. Figure
2\footnote{The regression model results are given in
  \ref{section:modelResult}. We ran another model where we used the
  self-reported acquisition time as a predictor with 3 levels 0-3:Early,
  4-7:Late1, and 7-above:Late2. However, our model did not converge
  properly and the credible intervals were very wide spread, yielding
  the results uninformative. We provide the model results as a plot in
  \ref{section:d}.} shows the posterior probability distributions where
the dot is the median estimate, the thick line is the 50\% credible
intervals, and the thin line is the 95\% credible intervals. There are 3
responses we are considering in this model. They are Classifier,
Nominal, and Zero Anaphora. Classifier response is the level that other
responses Nominal and Zero Anaphora are compared against.

@@

\hypertarget{regression-model-results-for-re-type}{%
\section{Regression model results for RE
Type}\label{regression-model-results-for-re-type}}

\label{section:modelResult}

\caption{\label{tab:unnamed-chunk-4}Regression model results for the RE Type (Nominal, Zero, Classifier) with the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (DoD, DoH)}

\begin{tabular}{lllllllll}
\toprule
Coefficients & \multicolumn{1}{c}{outerWidth} & \multicolumn{1}{c}{innerWidth} & \multicolumn{1}{c}{pointEstimate} & \multicolumn{1}{c}{ll} & \multicolumn{1}{c}{l} & \multicolumn{1}{c}{m} & \multicolumn{1}{c}{h} & \multicolumn{1}{c}{hh}\\
\midrule
Nom\_Intercept & 0.90 & 0.50 & median & 0.86 & 1.06 & 1.18 & 1.31 & 1.50\\
Zero\_Intercept & 0.90 & 0.50 & median & 0.11 & 0.26 & 0.37 & 0.46 & 0.59\\
Nom\_DoD & 0.90 & 0.50 & median & -0.67 & -0.31 & -0.06 & 0.18 & 0.55\\
Nom\_Intro & 0.90 & 0.50 & median & 3.34 & 3.56 & 3.72 & 3.88 & 4.13\\
Nom\_Maintain & 0.90 & 0.50 & median & -5.70 & -5.49 & -5.34 & -5.20 & -4.98\\
Nom\_DoD*Intro & 0.90 & 0.50 & median & -0.09 & 0.34 & 0.65 & 0.98 & 1.44\\
Nom\_DoD*Maintain & 0.90 & 0.50 & median & -1.86 & -1.45 & -1.16 & -0.88 & -0.46\\
Zero\_DoD & 0.90 & 0.50 & median & 0.45 & 0.72 & 0.89 & 1.10 & 1.40\\
Zero\_Intro & 0.90 & 0.50 & median & -4.77 & -4.22 & -3.87 & -3.55 & -3.13\\
Zero\_Maintain & 0.90 & 0.50 & median & 1.75 & 1.99 & 2.17 & 2.37 & 2.66\\
Zero\_DoD*Intro & 0.90 & 0.50 & median & 0.10 & 0.93 & 1.58 & 2.26 & 3.32\\
Zero\_DoD*Maintain & 0.90 & 0.50 & median & -1.95 & -1.38 & -0.99 & -0.64 & -0.17\\
\bottomrule
\end{tabular}

@@

\caption{\label{tab:unnamed-chunk-5}Model results for Nativeness with 3 levels}

@@

\leavevmode\vadjust pre{\hypertarget{ref-chamberlain_language_1999}{}}%
Chamberlain, Charlene, Morford, Jill P, Mayberry, Rachel I, 1999.
Language acquisition by eye. Psychology Press, Mahwah, N.J.

@@

\leavevmode\vadjust pre{\hypertarget{ref-czubek2017}{}}%
Czubek, Todd Alan, 2017. A comprehensive study of referring expressions
in ASL (PhD thesis).

@@

\leavevmode\vadjust pre{\hypertarget{ref-engberg-pedersen1993}{}}%
Engberg-Pedersen, Elisabeth, 1993. Space in danish sign language: The
semantics and morphosyntax of the use of space in a visual language.
Gallaudet University Press.

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-ministryofnationaleducation2020}{}}%
National Education, Ministry of, 2020.
\href{http://sgb.meb.gov.tr/meb_iys_dosyalar/2020_09/04144812_meb_istatistikleri_orgun_\%20egitim_2019_2020.pdf}{Milli
e{ğ}itim istatistikleri, örgün e{ğ}itim 2019/{'}20 {[}national education
statistics, formal education 2019/{'}20{]}}.
